A Makefile for Developing Containerized \LaTeX\ Technical Documents

### 2022-10-06 ms.tex
fix

\begin{document}
\DTLloaddb{keys-values}{artifacts/keys-values.csv}

@@

\begin{figure}[!t]
	\centering
	\begin{tikzpicture}[%
			grow via three points={one child at (0.5,-0.7) and two children at (0.5,-0.7) and (0.5,-1.4)},
			edge from parent path={(\tikzparentnode.south) |- (\tikzchildnode.west)}]
			\node[draw](document){document/}
			child{node[draw, anchor=west](makefile){Makefile}}
			child{node[draw, anchor=west, fill=red!20](tex){ms.tex}}
			child{node[draw, anchor=west, fill=red!20](bib){ms.bib}}
			child{node[draw, anchor=west, fill=blue!20](py){main.py}}
			child{node[draw, anchor=west, fill=blue!20](containerfile){Containerfile}}
			child{node[draw, anchor=west, fill=blue!20](requirements){requirements.txt}};
			\begin{scope}[on background layer]
				\node[draw, rounded corners, dashed, fill=gray!20, inner sep=4pt, fit=(document) (requirements)]{};
				\node[draw, rounded corners, dashed, fill=gray!80, inner sep=2pt, fit=(py) (requirements)](code){};
				\node[draw, rounded corners, dashed, fill=gray!80, inner sep=2pt, fit=(tex) (bib)](document){};
			\end{scope}
			\node[draw, right=1.5cm of code, fill=green!20] (results) {artifacts/results};
			\path[draw, ->] (code) -- (results);
			\node[draw, right=4cm of document, fill=green!20] (pdf) {artifacts/ms.pdf};
			\path[draw, ->] (results) |- (pdf);
			\path[draw, ->] (document) -- (pdf);
	\end{tikzpicture}
	\caption{The files of a repository that uses the Makefile is depicted in the light gray background and the arrows denote the containerized data flow towards generating the \textit{document}.
	Green indicates temporary cache or generated files, red \textit{\LaTeX\ code} files, blue \textit{results code} files and within the dark gray background, files that are developed by the \textit{author}.}\label{fig:repositoryfiles}
\end{figure}

@@

The pseudoalgorithm of the Makefile is as follows:
\begin{lstlisting}[language=bash, style=lststyle, caption={Pseudoalgorithm of the Makefile.}]
artifacts/$(document_pdf): $(latex_code) artifacts/$(results)
	run container latex

artifacts/$(results): $(results_code)
	build image $(results_image)
	run container $(results_image)

clean:
	rm -rf artifacts/
\end{lstlisting}

@@

The basic commands of the Makefile are the following:
\begin{lstlisting}[language=bash, style=lststyle, caption={Basic Makefile commands.}]
make                # Generate draft artifacts (fast).\ 
make FULL=1         # Generate full artifacts (slow).\ 
make clean          # Remove artifacts/ directory.
\end{lstlisting}

@@

The only requirements that the Makefile imposes on \textit{code} is the presence of \textit{results code} that generates the draft and full versions of the \textit{document} and the presence of \textit{results code} and \textit{\LaTeX\ code} for saving/loading the \textit{results} to/from \textit{artifacts/} respectively.
An example of a \textit{results code} that generates the draft version \textit{document} in the field of neural networks would be a conditional that sets the number of epochs and training/validation/test samples to a low value.

@@

A useful \LaTeX\ package for automatically presenting variables from \textit{results} in \textit{\LaTeX\ code} is \textit{datatool} with the following example use:
\begin{lstlisting}[language=TeX, style=lststyle, caption={\LaTeX\ datatool example of loading a file that contains pairs of keys and values (artifacts/keys-values.csv) generated by a \textit{results code} and getting the value of a key named lr.}]
\DTLloaddb{keys-values}{artifacts/keys-values.csv}
\DTLfetch{keys-values}{key}{lr}{value}
\end{lstlisting}

When using Python, a useful function for \textit{results code} is \textit{pandas.DataFrame.to\_latex} which automatically converts a dataframe table to a \LaTeX\ table:
\begin{lstlisting}[language=python, style=lststyle, caption={Convert Pandas DataFrame (df) to \LaTeX\ table (artifacts/metrics.tex) in \textit{results code}.}]
df = pd.DataFrame(table)
df.to_latex("artifacts/metrics.tex", float_format="%.2f")
\end{lstlisting}

@@

Additionally, an \textit{author} can use the \textit{cmp} utility to compare \textit{documents} (generated in a different session, operating system or underlying hardware) byte by byte, and diffoscope for identifying sources of non-determinism as follows:
\begin{lstlisting}[language=bash, style=lststyle, caption={Test draft document version reproducibility. This can also be used as a test script when pushing or pull requesting to a remote repository.}]
# Generate the first draft document version.\ 
make
# Backup the first resulting pdf.\ 
cp artifacts/ms.pdf artifacts/ms-previous.pdf
# Update the modification timestamp of main.py.\ 
touch main.py
# Generate the second draft document version.\ 
make
# Compare draft document versions byte by byte.\ 
cmp artifacts/ms.pdf artifacts/ms-previous.pdf
# Compare draft document versions
# in a human readable form.\ 
diffoscope artifacts/ms.pdf artifacts/ms-previous.pdf
\end{lstlisting}

@@

\begin{figure}[!t]
	\subfloat{\includegraphics[width=0.5\columnwidth]{artifacts/MNIST-loss.pdf}}
	\subfloat{\includegraphics[width=0.5\columnwidth]{artifacts/FashionMNIST-loss.pdf}}
	\\
	\subfloat{\includegraphics[width=0.5\columnwidth]{artifacts/KMNIST-loss.pdf}}
	\subfloat{\includegraphics[width=0.5\columnwidth]{artifacts/QMNIST-loss.pdf}}
	\caption{Comparison of the train and validation loss of MobilenetV2 with ReLU, ReLU6 and SiLU for all datasets.}\label{fig:loss}
\end{figure}

\begin{figure}[!t]
	\subfloat{\includegraphics[width=0.5\columnwidth]{artifacts/MNIST-ReLU-kernels.pdf}}
	\subfloat{\includegraphics[width=0.5\columnwidth]{artifacts/FashionMNIST-ReLU-kernels.pdf}}
	\\
	\subfloat{\includegraphics[width=0.5\columnwidth]{artifacts/KMNIST-ReLU-kernels.pdf}}
	\subfloat{\includegraphics[width=0.5\columnwidth]{artifacts/QMNIST-ReLU-kernels.pdf}}
	\caption{The first 25 kernels of the first convolutional layer of MobilenetV2 with ReLU for MNIST (upper-left), FashionMNIST (upper-right), KMNIST (lower-left) and QMNIST (lower-right).}\label{fig:kernels}
\end{figure}

\begin{table}[ht]
	\centering
	\caption{MobilenetV2 test dataset accuracies.}\label{table:table}
	\setlength\tabcolsep{4pt}
	\input{artifacts/metrics.tex}
\end{table}

>>>>

\begin{document}
\DTLloaddb{keys-values}{python/bin/keys-values.csv}

@@

\begin{figure}[!t]
	\centering
	\begin{tikzpicture}[%
			grow via three points={one child at (0.5,-0.7) and two children at (0.5,-0.7) and (0.5,-1.4)},
			edge from parent path={(\tikzparentnode.south) |- (\tikzchildnode.west)}]
			\node[draw](document){document/}
			child{node[draw, anchor=west](makefile){Makefile}}
			child{node[draw, anchor=west, fill=red!20](tex){ms.tex}}
			child{node[draw, anchor=west, fill=red!20](bib){ms.bib}}
			child{node[draw, anchor=west, fill=blue!20](py){main.py}}
			child{node[draw, anchor=west, fill=blue!20](containerfile){Containerfile}}
			child{node[draw, anchor=west, fill=blue!20](requirements){requirements.txt}};
			\begin{scope}[on background layer]
				\node[draw, rounded corners, dashed, fill=gray!20, inner sep=4pt, fit=(document) (requirements)]{};
				\node[draw, rounded corners, dashed, fill=gray!80, inner sep=2pt, fit=(py) (requirements)](code){};
				\node[draw, rounded corners, dashed, fill=gray!80, inner sep=2pt, fit=(tex) (bib)](document){};
			\end{scope}
			\node[draw, right=1.5cm of code, fill=green!20] (results) {python/bin/results};
			\path[draw, ->] (code) -- (results);
			\node[draw, right=4cm of document, fill=green!20] (pdf) {python/bin/ms.pdf};
			\path[draw, ->] (results) |- (pdf);
			\path[draw, ->] (document) -- (pdf);
	\end{tikzpicture}
	\caption{The files of a repository that uses the Makefile is depicted in the light gray background and the arrows denote the containerized data flow towards generating the \textit{document}.
	Green indicates temporary cache or generated files, red \textit{\LaTeX\ code} files, blue \textit{results code} files and within the dark gray background, files that are developed by the \textit{author}.}\label{fig:repositoryfiles}
\end{figure}

@@

The pseudoalgorithm of the Makefile is as follows:
\begin{lstlisting}[language=bash, style=lststyle, caption={Pseudoalgorithm of the Makefile.}]
bin/$(document_pdf): $(latex_code) bin/$(results)
	run container latex

bin/$(results): $(results_code)
	build image $(results_image)
	run container $(results_image)

clean:
	rm -rf bin/
\end{lstlisting}

@@

The basic commands of the Makefile are the following:
\begin{lstlisting}[language=bash, style=lststyle, caption={Basic Makefile commands.}]
make                # Generate draft bin (fast).\ 
make FULL=1         # Generate full bin (slow).\ 
make clean          # Remove bin/ directory.
\end{lstlisting}

@@

The only requirements that the Makefile imposes on \textit{code} is the presence of \textit{results code} that generates the draft and full versions of the \textit{document} and the presence of \textit{results code} and \textit{\LaTeX\ code} for saving/loading the \textit{results} to/from \textit{bin/} respectively.
An example of a \textit{results code} that generates the draft version \textit{document} in the field of neural networks would be a conditional that sets the number of epochs and training/validation/test samples to a low value.

@@

A useful \LaTeX\ package for automatically presenting variables from \textit{results} in \textit{\LaTeX\ code} is \textit{datatool} with the following example use:
\begin{lstlisting}[language=TeX, style=lststyle, caption={\LaTeX\ datatool example of loading a file that contains pairs of keys and values (bin/keys-values.csv) generated by a \textit{results code} and getting the value of a key named lr.}]
\DTLloaddb{keys-values}{bin/keys-values.csv}
\DTLfetch{keys-values}{key}{lr}{value}
\end{lstlisting}

When using Python, a useful function for \textit{results code} is \textit{pandas.DataFrame.to\_latex} which automatically converts a dataframe table to a \LaTeX\ table:
\begin{lstlisting}[language=python, style=lststyle, caption={Convert Pandas DataFrame (df) to \LaTeX\ table (bin/metrics.tex) in \textit{results code}.}]
df = pd.DataFrame(table)
df.to_latex("bin/metrics.tex", float_format="%.2f")
\end{lstlisting}

@@

Additionally, an \textit{author} can use the \textit{cmp} utility to compare \textit{documents} (generated in a different session, operating system or underlying hardware) byte by byte, and diffoscope for identifying sources of non-determinism as follows:
\begin{lstlisting}[language=bash, style=lststyle, caption={Test draft document version reproducibility. This can also be used as a test script when pushing or pull requesting to a remote repository.}]
# Generate the first draft document version.\ 
make
# Backup the first resulting pdf.\ 
cp bin/ms.pdf bin/ms-previous.pdf
# Update the modification timestamp of main.py.\ 
touch main.py
# Generate the second draft document version.\ 
make
# Compare draft document versions byte by byte.\ 
cmp bin/ms.pdf bin/ms-previous.pdf
# Compare draft document versions
# in a human readable form.\ 
diffoscope bin/ms.pdf bin/ms-previous.pdf
\end{lstlisting}

@@

\begin{figure}[!t]
	\subfloat{\includegraphics[width=0.5\columnwidth]{python/bin/MNIST-loss.pdf}}
	\subfloat{\includegraphics[width=0.5\columnwidth]{python/bin/FashionMNIST-loss.pdf}}
	\\
	\subfloat{\includegraphics[width=0.5\columnwidth]{python/bin/KMNIST-loss.pdf}}
	\subfloat{\includegraphics[width=0.5\columnwidth]{python/bin/QMNIST-loss.pdf}}
	\caption{Comparison of the train and validation loss of MobilenetV2 with ReLU, ReLU6 and SiLU for all datasets.}\label{fig:loss}
\end{figure}

\begin{figure}[!t]
	\subfloat{\includegraphics[width=0.5\columnwidth]{python/bin/MNIST-ReLU-kernels.pdf}}
	\subfloat{\includegraphics[width=0.5\columnwidth]{python/bin/FashionMNIST-ReLU-kernels.pdf}}
	\\
	\subfloat{\includegraphics[width=0.5\columnwidth]{python/bin/KMNIST-ReLU-kernels.pdf}}
	\subfloat{\includegraphics[width=0.5\columnwidth]{python/bin/QMNIST-ReLU-kernels.pdf}}
	\caption{The first 25 kernels of the first convolutional layer of MobilenetV2 with ReLU for MNIST (upper-left), FashionMNIST (upper-right), KMNIST (lower-left) and QMNIST (lower-right).}\label{fig:kernels}
\end{figure}

\begin{table}[ht]
	\centering
	\caption{MobilenetV2 test dataset accuracies.}\label{table:table}
	\setlength\tabcolsep{4pt}
	\input{python/bin/metrics.tex}
\end{table}
